---
title: "Test run time and memory usage of imputation accuracy"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
author: "Stefan McKinnon Hoj-Edwards"
output: 
  html_document: 
    theme: spacelab
    toc: yes
    toc_float: true
---

This vignette is for testing memory performance of Siccuracy's `imputation_accuracy` function using 
`fast=TRUE` and `fast=FALSE`. When `fast=TRUE`, neither matrix is stored in memory, whereas when `fast=FALSE`,
the true genotype file is stored in memory for matching row IDs.

Further more, using `standardization=TRUE` forces the methods to read the true genotype file *twice* in order to calculate means and standard deviations of true genotypes.

In order to estimate memory usage, the calculation is launched in a subprocess, while using Gregor Gorjanc's [cpumemlog](https://github.com/gregorgorjanc/cpumemlog) to monitor the usage of the subprocess.

```{r setup, include=FALSE}
library(Siccuracy)

BASEDIR <- getwd()
TMPDIR <- Sys.getenv('TMPDIR',tempdir())
TASKID <- Sys.getenv('SGE_TASK_ID', 'undefined')
suppressWarnings(TASKID <- as.integer(TASKID))
if (is.na(TASKID)) TASKID <- 0

.dot <- function(x) {
  f <- function(..., relative=FALSE) {
    if (relative) {
      file.path(file.path(x, c(...)))
    } else {
      file.path(BASEDIR, x, c(...))
    }
  } 
  null <- dir.create(f(''), FALSE, TRUE)
  return(f)
}
.cached <- .dot('cache_imputation_accuracy_memory_test')
.results <- .dot('imputation_accuracy_memory_test_results')

knitr::opts_knit$set(root.dir=TMPDIR)
knitr::opts_chunk$set(echo = TRUE, cache=TRUE, autodep=TRUE, 
                      cache.path = .cached('/'), 
                      fig.path = .results('/', relative=TRUE), 
                      fig.width=10, fig.height=7)


```

# Templates and preparation

Contents of `launcher.sh`:
```{r cache=FALSE,engine='cat',engine.opts=list(file='launcher.sh',lang='bash')}
Rscript --no-restore prepare.R $1 $2 $3 $4
Rscript --no-restore test.R $1 $2 $3 $4 &
pid=$!
./cpumemlog.sh $pid -o=$5 -t=1
wait
```

Contents of `prepare.R`:
```{r cache=FALSE,engine='cat',engine.opts=list(file='prepare.R',lang='R')}
library(Siccuracy)

args <- commandArgs(TRUE)
n <- as.integer(args[1])
m <- as.integer(args[2])
fast <- as.logical(args[3])
standardized <- as.logical(args[4])

true <- Siccuracy:::make.true(n, m)
impu <- Siccuracy:::make.imputed(true)

fn1 <- 'true.txt'
fn2 <- 'impu.txt'

unlink(c(fn1,fn2))

write.snps(true, fn1)
write.snps(impu, fn1)
```

Contents of `test.R`:

```{r cache=FALSE,engine='cat',engine.opts=list(file='test.R',lang='R')}
library(Siccuracy)

args <- commandArgs(TRUE)
n <- as.integer(args[1])
m <- as.integer(args[2])
fast <- as.logical(args[3])
standardized <- as.logical(args[4])

res <- imputation_accuracy('true.txt','impu.txt', fast = fast, standardized = standardized)
```

```{r cache=FALSE}
file.copy(file.path(BASEDIR, '../tools/cpumemlog/cpumemlog.sh'), TMPDIR, overwrite = TRUE)
system2('chmod', c('u+x', 'cpumemlog.sh','launcher.sh'))

launch <- function(n, m, fast, standardized, fn=NULL) {
  fast <- as.character(fast)
  standardized <- as.character(standardized)
  if (is.null(fn)) fn <- .results(paste0(paste('cpumemlog',n,m,fast,standardized,sep='_'), '.txt'))
  st <- system2('./launcher.sh', args=c(n,m,fast,standardized,fn), env=paste0('R_LIBS=', paste(.libPaths(), collapse=':')))
  if (st != 0) return(NULL)
  fn
}

ns <- c(100, 500, 1000, 2000, 5000, 1e5, 5e5, 1e6)
ms <- c(1000, 2000, 5000, 1e5, 5e5, 1e6)
```

## Non-fast, no standardisation

```{r nofast_nostandard,eval=TASKID==0|TASKID==1}
for (n in c(10,100,500,1000,2000,5000)) {
  for (m in c(100,1000,2000,5000,8000,10000)) {
    launch(n, m, fast=FALSE, standardized=FALSE)
  }
}
```


# Code preamble

The following code block is executed in the very beginning of the script.
```{r setup,eval=FALSE}
```